### 2014-02-14 paper.tex
Add basic paper.

%% bare_conf.tex
%% V1.4
%% 2012/12/27
%% by Michael Shell
%% See:
%% http://www.michaelshell.org/
%% for current contact information.
%%
%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.8 or later) with an IEEE conference paper.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/tex-archive/macros/latex/contrib/IEEEtran/
%% and
%% http://www.ieee.org/

%%*************************************************************************
%% Legal Notice:
%% This code is offered as-is without any warranty either expressed or
%% implied; without even the implied warranty of MERCHANTABILITY or
%% FITNESS FOR A PARTICULAR PURPOSE! 
%% User assumes all risk.
%% In no event shall IEEE or any contributor to this code be liable for
%% any damages or losses, including, but not limited to, incidental,
%% consequential, or any other damages, resulting from the use or misuse
%% of any information contained here.
%%
%% All comments are the opinions of their respective authors and are not
%% necessarily endorsed by the IEEE.
%%
%% This work is distributed under the LaTeX Project Public License (LPPL)
%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
%% distributed and modified. A copy of the LPPL, version 1.3, is included
%% in the base LaTeX documentation of all distributions of LaTeX released
%% 2003/12/01 or later.
%% Retain all contribution notices and credits.
%% ** Modified files should be clearly indicated as such, including  **
%% ** renaming them and changing author support contact information. **
%%
%% File list of work: IEEEtran.cls, IEEEtran_HOWTO.pdf, bare_adv.tex,
%%                    bare_conf.tex, bare_jrnl.tex, bare_jrnl_compsoc.tex,
%%                    bare_jrnl_transmag.tex
%%*************************************************************************

% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
% *** with production work. IEEE's font choices can trigger bugs that do  ***
% *** not appear when using other class files.                            ***
% The testflow support page is at:
% http://www.michaelshell.org/tex/testflow/

% Note that the a4paper option is mainly intended so that authors in
% countries using A4 can easily print to A4 and see how their papers will
% look in print - the typesetting of the document will not typically be
% affected with changes in paper size (but the bottom and side margins will).
% Use the testflow package mentioned above to verify correct handling of
% both paper sizes by the user's LaTeX system.
%
% Also note that the "draftcls" or "draftclsnofoot", not "draft", option
% should be used if it is desired that the figures are to be displayed in
% draft mode.
%
\documentclass[conference]{IEEEtran}
% Add the compsoc option for Computer Society conferences.
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[conference]{../sty/IEEEtran}

% Some very useful LaTeX packages include:
% (uncomment the ones you want to load)

% *** MISC UTILITY PACKAGES ***
%
%\usepackage{ifpdf}
% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
% compilation based on whether the output is pdf or dvi.
% usage:
% \ifpdf
%   % pdf code
% \else
%   % dvi code
% \fi
% The latest version of ifpdf.sty can be obtained from:
% http://www.ctan.org/tex-archive/macros/latex/contrib/oberdiek/
% Also, note that IEEEtran.cls V1.7 and later provides a builtin
% \ifCLASSINFOpdf conditional that works the same way.
% When switching from latex to pdflatex and vice-versa, the compiler may
% have to be run twice to clear warning/error messages.

% *** CITATION PACKAGES ***
%
%\usepackage{cite}
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off
% such as if a citation ever needs to be enclosed in parenthesis.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 4.0 (2003-05-27) and later if using hyperref.sty. cite.sty does
% not currently provide for hyperlinked citations.
% The latest version can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/cite/
% The documentation is contained in the cite.sty file itself.

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
  % \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation
% can be obtained at: 
% http://www.ctan.org/tex-archive/macros/latex/required/graphics/
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found at:
% http://www.ctan.org/tex-archive/info/epslatex/
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

% *** MATH PACKAGES ***
%
%\usepackage[cmex10]{amsmath}
% A popular package from the American Mathematical Society that provides
% many useful and powerful commands for dealing with mathematics. If using
% it, be sure to load this package with the cmex10 option to ensure that
% only type 1 fonts will utilized at all point sizes. Without this option,
% it is possible that some math symbols, particularly those within
% footnotes, will be rendered in bitmap form which will result in a
% document that can not be IEEE Xplore compliant!
%
% Also, note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
%\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/amslatex/math/

% *** SPECIALIZED LIST PACKAGES ***
%
%\usepackage{algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithms/
% There is also a support site at:
% http://algorithms.berlios.de/index.html
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithmicx/

% *** ALIGNMENT PACKAGES ***
%
%\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/tools/

% IEEEtran contains the IEEEeqnarray family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

% *** SUBFIGURE PACKAGES ***
%\ifCLASSOPTIONcompsoc
%  \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
%\else
%  \usepackage[caption=false,font=footnotesize]{subfig}
%\fi
% subfig.sty, written by Steven Douglas Cochran, is the modern replacement
% for subfigure.sty, the latter of which is no longer maintained and is
% incompatible with some LaTeX packages including fixltx2e. However,
% subfig.sty requires and automatically loads Axel Sommerfeldt's caption.sty
% which will override IEEEtran.cls' handling of captions and this will result
% in non-IEEE style figure/table captions. To prevent this problem, be sure
% and invoke subfig.sty's "caption=false" package option (available since
% subfig.sty version 1.3, 2005/06/28) as this is will preserve IEEEtran.cls
% handling of captions.
% Note that the Computer Society format requires a larger sans serif font
% than the serif footnote size font used in traditional IEEE formatting
% and thus the need to invoke different subfig.sty package options depending
% on whether compsoc mode has been enabled.
%
% The latest version and documentation of subfig.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/subfig/

% *** FLOAT PACKAGES ***
%
%\usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure. The latest version and documentation can be found at:
% http://www.ctan.org/tex-archive/macros/latex/base/

%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/sttools/
% Do not use the stfloats baselinefloat ability as IEEE does not allow
% \baselineskip to stretch. Authors submitting work to the IEEE should note
% that IEEE rarely uses double column equations and that authors should try
% to avoid such use. Do not be tempted to use the cuted.sty or midfloat.sty
% packages (also by Sigitas Tolusis) as IEEE does not format its papers in
% such ways.
% Do not attempt to use stfloats with fixltx2e as they are incompatible.
% Instead, use Morten Hogholm'a dblfloatfix which combines the features
% of both fixltx2e and stfloats:
%
% \usepackage{dblfloatfix}
% The latest version can be found at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/dblfloatfix/

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
%\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/url/
% Basically, \url{my_url_here}.

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

% correct bad hyphenation here

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
% Do not put math or special symbols in the title.
\title{Something to do with Image Processing}

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{\IEEEauthorblockN{Alexander David Brown}
\IEEEauthorblockA{Computer Science,\\
Aberystwyth University,\\
Penglais,
Aberystwyth,\\
Ceredigion,\\
Wales SY23 3DB\\
Email: adb9@aber.ac.uk}
}

% conference papers do not typically use \thanks and this command
% is locked out in conference mode. If really needed, such as for
% the acknowledgment of grants, issue a \IEEEoverridecommandlockouts
% after \documentclass

% for over three affiliations, or if they all won't fit within the width
% of the page, use this alternative format:
% 
%\author{\IEEEauthorblockN{Michael Shell\IEEEauthorrefmark{1},
%Homer Simpson\IEEEauthorrefmark{2},
%James Kirk\IEEEauthorrefmark{3}, 
%Montgomery Scott\IEEEauthorrefmark{3} and
%Eldon Tyrell\IEEEauthorrefmark{4}}
%\IEEEauthorblockA{\IEEEauthorrefmark{1}School of Electrical and Computer Engineering\\
%Georgia Institute of Technology,
%Atlanta, Georgia 30332--0250\\ Email: see http://www.michaelshell.org/contact.html}
%\IEEEauthorblockA{\IEEEauthorrefmark{2}Twentieth Century Fox, Springfield, USA\\
%Email: homer@thesimpsons.com}
%\IEEEauthorblockA{\IEEEauthorrefmark{3}Starfleet Academy, San Francisco, California 96678-2391\\
%Telephone: (800) 555--1212, Fax: (888) 555--1212}
%\IEEEauthorblockA{\IEEEauthorrefmark{4}Tyrell Inc., 123 Replicant Street, Los Angeles, California 90210--4321}}

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

% make the title area
\maketitle

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
The abstract goes here.
\end{abstract}

% no keywords

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

\section{Introduction}
% no \IEEEPARstart

\section{Conclusion}
The conclusion goes here.

% conference papers do not normally have an appendix

% use section* for acknowledgement

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
%\bibliography{bibliography}

% that's all folks
\end{document}

### 2014-03-31 paper.tex
Add some content to the survey paper.

@@
 % Do not put math or special symbols in the title.
-\title{Something to do with Image Processing}
+\title{Digital Analysis of Paintings}
 
@@
 \begin{abstract}
-The abstract goes here.
+%TODO Write this once the paper is complete
 \end{abstract}
@@
 \section{Introduction}
-% no \IEEEPARstart
+Digital image processing is a field which encourages the crossover of different
+scientific disciplines, often biological research and computer vision align to
+automate the collection and analysis of plant growth. However, art and
+computing are fields which, at first glance, have very little in common.
+
+But a deeper look unveils a plethora of different avenues from detecting
+forgeries to being able to date an artists work within a catalogue of their
+known pieces\cite{brown2013can}.
+
+This survey paper will unearth some of the techniques which are used to perform
+such analysis and evaluate their application in real world problems.
+
+%TODO this still need some expansion.
+
+\section{Image Processing}
+A typical digital image is represented as a matrix of pixels where each pixel
+contains information about the colour it represents. This colour can be
+described using several different representations and although each
+representation will equate to the same colour in the rendered image, the
+information stored can be radically different.
+
+This representation of an image provides several key pieces of information,
+including:
+
+\begin{itemize}
+\item colour space of the image
+\item the relationship between pixels
+\end{itemize}
+
+The digital processing of images is normally done through the application of
+functions or filters to the matrix to gather specific information relating to
+the desired result. A simple example of this would be averaging the colours of
+an image to get a basic idea of the colours in an image.
+
+More advanced techniques involve performing several of these operations in
+sequence to provide a more complex view of the image.
+
+There are many sections of image processing, each with their different purposes
+this report will particularly focus on those relating to colour space; edge
+detection and orientation for texture analysis and segmentation and regions of
+interest for brush stroke analysis.
+
+% TODO write more here too
+
+\section{Colour Space Analysis}
+
+\section{Texture Analysis}
+
+\section{Brush Stroke Analysis}
+
+\section{Application of Image Processing to Artwork}
+
+\subsection{Dating Artists Work}
 
 \section{Conclusion}
-The conclusion goes here.
 
@@
 % argument is your BibTeX string definitions and bibliography database(s)
-%\bibliography{bibliography}
+\bibliography{bibliography}
 

### 2014-04-22 paper.tex
Add in potential titles.

--- 2014-02-14 paper.tex
@@
 %%*************************************************************************
-
-% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
-% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
-% *** with production work. IEEE's font choices can trigger bugs that do  ***
-% *** not appear when using other class files.                            ***
-% The testflow support page is at:
-% http://www.michaelshell.org/tex/testflow/
-
-% Note that the a4paper option is mainly intended so that authors in
-% countries using A4 can easily print to A4 and see how their papers will
-% look in print - the typesetting of the document will not typically be
-% affected with changes in paper size (but the bottom and side margins will).
-% Use the testflow package mentioned above to verify correct handling of
-% both paper sizes by the user's LaTeX system.
-%
-% Also note that the "draftcls" or "draftclsnofoot", not "draft", option
-% should be used if it is desired that the figures are to be displayed in
-% draft mode.
-%
 \documentclass[conference]{IEEEtran}
 % Add the compsoc option for Computer Society conferences.
-%
-% If IEEEtran.cls has not been installed into the LaTeX system files,
-% manually specify the path to it like:
-% \documentclass[conference]{../sty/IEEEtran}
 
-% Some very useful LaTeX packages include:
-% (uncomment the ones you want to load)
+\ifCLASSINFOpdf
+  \usepackage[pdftex]{graphicx}
+  \graphicspath{{../img/}}
+  \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
+\else
+  \usepackage[dvips]{graphicx}
+  \graphicspath{{../img/}}
+  \DeclareGraphicsExtensions{.eps}
+\fi
 
-% *** MISC UTILITY PACKAGES ***
-%
-%\usepackage{ifpdf}
-% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
-% compilation based on whether the output is pdf or dvi.
-% usage:
-% \ifpdf
-%   % pdf code
-% \else
-%   % dvi code
-% \fi
-% The latest version of ifpdf.sty can be obtained from:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/oberdiek/
-% Also, note that IEEEtran.cls V1.7 and later provides a builtin
-% \ifCLASSINFOpdf conditional that works the same way.
-% When switching from latex to pdflatex and vice-versa, the compiler may
-% have to be run twice to clear warning/error messages.
-
-% *** CITATION PACKAGES ***
-%
-%\usepackage{cite}
-% cite.sty was written by Donald Arseneau
-% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
-% \cite{} output to follow that of IEEE. Loading the cite package will
-% result in citation numbers being automatically sorted and properly
-% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
-% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
-% \cite will automatically add leading space, if needed. Use cite.sty's
-% noadjust option (cite.sty V3.8 and later) if you want to turn this off
-% such as if a citation ever needs to be enclosed in parenthesis.
-% cite.sty is already installed on most LaTeX systems. Be sure and use
-% version 4.0 (2003-05-27) and later if using hyperref.sty. cite.sty does
-% not currently provide for hyperlinked citations.
-% The latest version can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/cite/
-% The documentation is contained in the cite.sty file itself.
-
-% *** GRAPHICS RELATED PACKAGES ***
-%
-\ifCLASSINFOpdf
-  % \usepackage[pdftex]{graphicx}
-  % declare the path(s) where your graphic files are
-  % \graphicspath{{../pdf/}{../jpeg/}}
-  % and their extensions so you won't have to specify these with
-  % every instance of \includegraphics
-  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
-\else
-  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
-  % will default to the driver specified in the system graphics.cfg if no
-  % driver is specified.
-  % \usepackage[dvips]{graphicx}
-  % declare the path(s) where your graphic files are
-  % \graphicspath{{../eps/}}
-  % and their extensions so you won't have to specify these with
-  % every instance of \includegraphics
-  % \DeclareGraphicsExtensions{.eps}
-\fi
-% graphicx was written by David Carlisle and Sebastian Rahtz. It is
-% required if you want graphics, photos, etc. graphicx.sty is already
-% installed on most LaTeX systems. The latest version and documentation
-% can be obtained at: 
-% http://www.ctan.org/tex-archive/macros/latex/required/graphics/
-% Another good source of documentation is "Using Imported Graphics in
-% LaTeX2e" by Keith Reckdahl which can be found at:
-% http://www.ctan.org/tex-archive/info/epslatex/
-%
-% latex, and pdflatex in dvi mode, support graphics in encapsulated
-% postscript (.eps) format. pdflatex in pdf mode supports graphics
-% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
-% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
-% not a bitmapped formats (.jpeg, .png). IEEE frowns on bitmapped formats
-% which can result in "jaggedy"/blurry rendering of lines and letters as
-% well as large increases in file sizes.
-%
-% You can find documentation about the pdfTeX application at:
-% http://www.tug.org/applications/pdftex
-
-% *** MATH PACKAGES ***
-%
-%\usepackage[cmex10]{amsmath}
-% A popular package from the American Mathematical Society that provides
-% many useful and powerful commands for dealing with mathematics. If using
-% it, be sure to load this package with the cmex10 option to ensure that
-% only type 1 fonts will utilized at all point sizes. Without this option,
-% it is possible that some math symbols, particularly those within
-% footnotes, will be rendered in bitmap form which will result in a
-% document that can not be IEEE Xplore compliant!
-%
-% Also, note that the amsmath package sets \interdisplaylinepenalty to 10000
-% thus preventing page breaks from occurring within multiline equations. Use:
-%\interdisplaylinepenalty=2500
-% after loading amsmath to restore such page breaks as IEEEtran.cls normally
-% does. amsmath.sty is already installed on most LaTeX systems. The latest
-% version and documentation can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/required/amslatex/math/
-
-% *** SPECIALIZED LIST PACKAGES ***
-%
-%\usepackage{algorithmic}
-% algorithmic.sty was written by Peter Williams and Rogerio Brito.
-% This package provides an algorithmic environment fo describing algorithms.
-% You can use the algorithmic environment in-text or within a figure
-% environment to provide for a floating algorithm. Do NOT use the algorithm
-% floating environment provided by algorithm.sty (by the same authors) or
-% algorithm2e.sty (by Christophe Fiorio) as IEEE does not use dedicated
-% algorithm float types and packages that provide these will not provide
-% correct IEEE style captions. The latest version and documentation of
-% algorithmic.sty can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithms/
-% There is also a support site at:
-% http://algorithms.berlios.de/index.html
-% Also of interest may be the (relatively newer and more customizable)
-% algorithmicx.sty package by Szasz Janos:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithmicx/
+\usepackage[cmex10]{amsmath}
 
@@
 
-% *** PDF, URL AND HYPERLINK PACKAGES ***
-%
-%\usepackage{url}
-% url.sty was written by Donald Arseneau. It provides better support for
-% handling and breaking URLs. url.sty is already installed on most LaTeX
-% systems. The latest version and documentation can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/url/
-% Basically, \url{my_url_here}.
-
-% *** Do not adjust lengths that control margins, column widths, etc. ***
-% *** Do not use packages that alter fonts (such as pslatex).         ***
-% There should be no need to do such things with IEEEtran.cls V1.6 and later.
-% (Unless specifically asked to do so by the journal or conference you plan
-% to submit to, of course. )
+\usepackage{url}
 
@@
 \begin{document}
-%
-% paper title
-% can use linebreaks \\ within to get better formatting as desired
-% Do not put math or special symbols in the title.
-\title{Something to do with Image Processing}
 
-% author names and affiliations
-% use a multiple column layout for up to three different
-% affiliations
+\title{Digital Analysis of Paintings}
+
 \author{\IEEEauthorblockN{Alexander David Brown}
@@
 
-% conference papers do not typically use \thanks and this command
-% is locked out in conference mode. If really needed, such as for
-% the acknowledgment of grants, issue a \IEEEoverridecommandlockouts
-% after \documentclass
-
-% for over three affiliations, or if they all won't fit within the width
-% of the page, use this alternative format:
-% 
-%\author{\IEEEauthorblockN{Michael Shell\IEEEauthorrefmark{1},
-%Homer Simpson\IEEEauthorrefmark{2},
-%James Kirk\IEEEauthorrefmark{3}, 
-%Montgomery Scott\IEEEauthorrefmark{3} and
-%Eldon Tyrell\IEEEauthorrefmark{4}}
-%\IEEEauthorblockA{\IEEEauthorrefmark{1}School of Electrical and Computer Engineering\\
-%Georgia Institute of Technology,
-%Atlanta, Georgia 30332--0250\\ Email: see http://www.michaelshell.org/contact.html}
-%\IEEEauthorblockA{\IEEEauthorrefmark{2}Twentieth Century Fox, Springfield, USA\\
-%Email: homer@thesimpsons.com}
-%\IEEEauthorblockA{\IEEEauthorrefmark{3}Starfleet Academy, San Francisco, California 96678-2391\\
-%Telephone: (800) 555--1212, Fax: (888) 555--1212}
-%\IEEEauthorblockA{\IEEEauthorrefmark{4}Tyrell Inc., 123 Replicant Street, Los Angeles, California 90210--4321}}
-
-% use for special paper notices
-%\IEEEspecialpapernotice{(Invited Paper)}
-
-% make the title area
 \maketitle
 
-% As a general rule, do not put math, special symbols or citations
-% in the abstract
 \begin{abstract}
-The abstract goes here.
+%TODO Write this once the paper is complete
 \end{abstract}
@@
 
-% For peer review papers, you can put extra information on the cover
-% page as needed:
-% \ifCLASSOPTIONpeerreview
-% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
-% \fi
-%
-% For peerreview papers, this IEEEtran command inserts a page break and
-% creates the second title. It will be ignored for other modes.
 \IEEEpeerreviewmaketitle
@@
 \section{Introduction}
-% no \IEEEPARstart
+Digital image processing is a field which encourages the crossover of different
+scientific disciplines, often biological research and computer vision align to
+automate the collection and analysis of plant growth. However, art and
+computing are fields which, at first glance, have very little in common.
 
-\section{Conclusion}
-The conclusion goes here.
+But a deeper look unveils a plethora of different avenues from detecting
+forgeries to being able to date an artists work within a catalogue of their
+known pieces.
+
+This survey paper will unearth some techniques which can be applied to the
+digital analysis of paintings, as well as existing research which has already
+applied computer vision to the field of art.
+
+\section{Colour Analysis}
+
+\subsection{Pigment Mapping}
+% Paper 3
+
+\subsection{Difference Visualisation}
+% Paper 6
+
+\section{Texture Analysis}
+Paintings are somewhat unlike the normal subject for image processing, whilst
+most images are either simple two dimensional images, or two dimensional slices
+of a three dimensional object. Paintings often thought of as two dimensional,
+but with many paint types these paintings become three dimensional.
+
+This aspect is lost in the digitisation of the painting; but there are still
+ways of analysing the texture of the image.
+
+\subsection{Steerable Filters}
+% Paper 12
+
+\subsection{Gabor Filters}
+% Paper 13
+
+\subsection{Histograms of Edge Orientated Gradients}
+% Paper 8
+
+\subsection{Multifractal Classification}
+% Paper 5
+
+\subsection{Texton-Based Analysis}
+% Paper 11
+
+\section{Statistical Analysis}
+
+\subsection{Stylistic Analysis}
+% Paper 9
+
+\subsection{Authentication of Artwork}
+% Paper 4
+
+\subsection{Dating an Artist's Work}
+% Paper AWESOME :D
+
+\section{Brushstroke Analysis}
+
+\subsection{Artistic Identification}
+% Paper 1
+
+\subsection{Rhythmic Brushstrokes}
+% Paper 2
 

### 2014-04-25 paper.tex
Add another paper into the mix :-)
Add introduction for colour analysis.
I've probably waffled on a bit too much about colour spaces and such forth, but
I think it's all for a good cause (i.e. to set the scene for that section, and
for the rest of the paper).

@@
 \usepackage{url}
+\usepackage{glossaries}
+
+\newacronym{RGB}{RGB}{Red, Green, Blue}
 
@@
 \section{Colour Analysis}
+Digital images are typically considered to be a matrix of pixels, where each
+pixel contains information about the colour of that place in the image.
+
+Colours can be represented in numerous different ways, but are typically three
+to four bytes; for example the \gls{RGB} colour space uses one byte for the
+levels of red, one for green and one for blue. The fourth byte is unlikely to
+be considered in artwork as it usually represents the transparency, known as
+the alpha channel, of the pixel.
+
+A single byte colour space purely focuses on the intensity of a pixel, in
+actuality this represents a grayscale image.
+
+Analysing colour is the base of all analysis techniques in image processing,
+the value(s) of a pixel with regards to its location and neighbours can be used
+to build up some very complex knowledge about the image. This section will deal
+specifically with the colours which an artist uses within their work, such as
+looking at the distribution of colours across a work, rather than using colour
+information to determine textures.
+
+\subsection{Colour Distribution}
+% Paper 7
 

### 2014-04-26 paper.tex
Write up paper 7.

@@
 
+\newacronym{APICSS}{APICSS}{Art Painting Image Color Semantics}
 \newacronym{RGB}{RGB}{Red, Green, Blue}
+\newacronym{HSL}{HSL}{Hue, Saturation, Luminance}
+\newacronym{HSV}{HSV}{Hue, Saturation, Value}
 
@@
 % Paper 7
+
+The distribution of colour across an image can give a surprising amount of
+information about that image, especially given that colour is very subjective
+to an individual.
+
+Ivonna, Stanchev and Dimitrov investigated the colour
+distribution within the works of over 100 artists for several different
+countries and periods, using a system named \gls{APICSS}\cite{ivanova2008analysis}.
+
+Existing systems already statistically analysed colours within an image and
+some even used classifiers, typically a naive Bayesian classifier as it fits
+well with statical methods, to perform extra analysis.
+
+Unlike these existing system, \gls{APICSS} focused primarily on the \gls{HSL}
+colour space as it is the closest representation to an artist's colour wheel.
+\gls{HSV} was also considered, but because the lightness is not symmetrical
+within the \gls{HSV} colour space, it is less useful for direct comparison.
+
+As with any system involving classification, a feature space is needed to
+represent an item within the data set. In the case of \gls{APICSS} the feature
+space is three dimensional, one for each of the values in the colour space.
+
+From this, a distance measure (in this case Euclidean distance) can be used to
+compare different items.
+
+\gls{APICSS} also take into account some of the metadata attached to each
+painting, such as the movement and sub-movement during which the painting was
+created. This allows for more statistical analyses to be performed on sub-sets
+of the data set.
+
+\gls{APICSS} appears to be a good system, which considers both metadata and
+analysis gained from the digitisation of artwork. The paper considers a wide
+range of artists and periods.
+
+However, the use of a lossy image format (JPEG) could potentially skew the
+results. The system itself is very basic, especially if it only considers
+\gls{HSL} and metadata. This is further compounded by only considering thirteen
+colours within the Hue (twelve separate colours and one achromatic).
+
+Despite having a decent sized data set overall, some of the categories have
+relatively small items within them.
+
+Overall \gls{APICSS} is a good application of existing research, but uses very
+basic analysis techniques on the images themselves.
 

### 2014-04-26 paper.tex
Add paper 3.
Up to around 1.25k words :-) possibly writing a bit too much for each paper,
but we'll see.

@@
 % Paper 3
+Colour analysis can also give a glimpse into the colour pigments the artist
+used to create the work. Using a database of known oil pigments and
+\textit{a priori} knowledge, it is possible to work out which pigments are used
+in a painting, and also to be able to show which parts of the painting are made
+up from which pigment\cite{zhao2008investigation}.
+
+This is done through a process known a multispectral imaging, the process of
+capturing a different frequencies to provide a different spectrum of light and
+enable more analysis.
+
+This technique is especially powerful for segmenting different pigments in an
+image as each pigment is made up of a different substance.
+
+Because this process requires both hardware and software it make a large
+different whether the materials a work is made up of are known. When known it
+can be possible to find the combinations of these materials at a given point
+and their concentrations.
+
+When the materials are unknown, extra processing is required to decide which
+materials the painting comprises of. This process uses spectral processing but
+also requires prior knowledge of the artist and art tools.
+
+In the experimental results this paper provides, the latter is achieved by a
+study of another of Van Gogh's works from the same time.
+
+Pigment mapping across every pixel is a costly operation in terms of memory and
+CPU time. This research makes a reasonable assumption that a similar colour
+will have been painted using similar pigments to reduce the amount of pigment
+mapping which occurs. This does lead to unwanted side-effects, but the only way
+to combat these is to increase the number of pixels which pixel-mapping.
+
+To reduce the number of pixels considered for pigment mapping the image is
+first partitioned into a number of smaller images to reduce processing time on
+the next steps
+
+Each of these partitioned regions is then split into a number of clusters based
+on the colours of the pixels. This was done using supervised and unsupervised
+algorithms. The unsupervised algorithm was very sensitive and had long running
+times whilst the supervised algorithm requires a human observer.
+
+Finally pixels within these clusters were then selected based on the similarity
+of pixels within the cluster, the clusters and partitioned images can then be
+stitched back together and rendered based on the pigment maps.
+
+This research could be incredibly useful if it could be applied to the analysis
+of artwork and the field of multispectral analysis has already had a lot of
+success in other fields.
+
+However, because this is both a hardware- and software-based system, it may not
+be something which can be used with regularity as it requires the physical
+presence of the paintings, which not all artwork analysis researchers have
+access to.
+
+It also requires a lot of a priori knowledge of the artist and calculating the
+materials a painting consists of may be a difficult, if not impossible task,
+depending on the artist.
+
+In conclusion, this technique can provide a very powerful analysis of a
+painting, but at the cost of requiring specialised hardware, the physical
+painting and expert knowledge about the artist, so is a very manual process.
+
+This technique could be improved by using a less naive clustering unsupervised
+algorithm to remove some of the need for human intervention, but the need for
+the physical painting will always be a limiting factor of this work.
 

### 2014-04-26 paper.tex
Add another paper.

@@
 % Paper 6
+Another approach to colour is to visualise the difference between two or more
+different images, this is particularly of use for comparing different versions
+of the same image - the original to a famous forgery, for example.
+
+Image fusion is the process of combining two different images and has recently
+been applied to artwork\cite{blazek13image}. In this research, a technique for
+showing the differences between two or more images which can show positive,
+negative and zero difference between the images, whilst maintaining regions of
+interest between the two images.
+
+The method for doing this uses the CIELAB colour space, which maps more
+faithfully to human vision than other colour spaces. However, because of the
+limitations displaying the images, they are converted to \gls{HSV} before they
+are displayed.
+
+A hue for both the positive and negative difference values is selected by the
+user as a vector, then the difference in the \gls{HSV} saturation of the two
+images is then computed into CIELAB, where it is an angle between the
+aforementioned colour vector.
+
+Finally this is converted into \gls{HSV}, where the Hue represents the sign of
+the computed image, the saturation the absolute distance and the value the
+average value of intensities of the compared image at that pixel.
+
+The authors claim this fusion improves the comprehension of the difference
+visualisation of images and that it makes identifying regions of interest
+faster and more precise, but provide no experimental results to prove this
+other than the fused image result of four different copies of the same artwork.
+
+This does keep the original image intact (albeit in grayscale with differences
+highlighted).
+
+This does only provide visualisation rather than analysis, but could potential
+be used as part of a larger system to build more complex analysis techniques.
+This might be especially useful to pick up difference between an original and a
+derivative work or forgery.
+
+At current, the images must be aligned and scaled manually, but this process
+could be automated trivially.
 
@@
 This aspect is lost in the digitisation of the painting; but there are still
-ways of analysing the texture of the image.
+ways of analysing the texture of the image by analysing the colour of the
+image. Quite often, this is as effective on a one dimensional colour space as
+it is on a three dimensional one.
+
+Filters can be passed over an image to gather basic information such as
+direction of lines within the paintings. More complex techniques for analysing
+texture involve using wavelets and considering nearby pixels.
 

### 2014-04-27 paper.tex
Add HOG paper details.

@@
 
+This research appears to be of very limited use, but when used within as a part
+of a larger system could provide some very interesting analysis which could be
+used to train classifiers to detect forgeries, etc. but this still has limited
+scope.
+
 \section{Texture Analysis}
@@
 % Paper 8
+
+A method for analysing the texture over an image to to create a histogram which
+contains the orientation of all the gradients over an image. This technique has
+been applied to the field of human detection\cite{dalal05histograms}, but is
+also useful when applied to the realm of digital analysis of artwork as well.
+
+A lot of the methodology for human detection involves the normalisation and
+classification, but does include some useful information about the practises of
+generating histograms of edge-orientated gradients. The gradients are computed
+using a variety of discrete derivative masks combined with degrees of Gaussian
+smoothing, although the results show that Gaussian smoothing and larger masks
+damage the performance.
+
+Each pixel is then binned according to a weighted vote from the mask, then each
+vote is accumulated into orientation bins in local regions, which could be
+rectangular or radial.
+
+These orientation bins were evenly spaced across $0-\pi$ unsigned bins or
+$i-2\pi$ signed bins. This is significant to note an orientation $\theta$ below
+$\pi$ is often thought of as equivalent to $\theta + \pi$ in image processing.
+
+Normalisation is then applied to reduce the effects of local contrast and is
+good for performance in the field of human detection. For artwork this may be
+less useful as these features are of importance.
+
+This research shows very good experimental results of large, popular data sets
+and beats all of the existing techniques on false positives (although no data
+for false negatives is shown).
+
+Although this research does not directly relate to the field of digital
+analysis of artwork, it has been used to generate some interesting
+analysis\cite{brown13can} although not all of the technique was applied in that
+research.
+
+The use of this research is the ability to store orientation data in
+histograms, which are very easy to process digitally.
+
+In conclusion, this technique provides a powerful analysis technique and the
+experimental results on several decent sized data sets to prove this. Although
+not all of the technique is useful or applicable to the digital analysis of
+paintings, the parts which are have been successfully used in research into
+artwork.
 

### 2014-04-29 paper.tex
More a good stab at a few of the papers.

@@
   \usepackage[pdftex]{graphicx}
-  \graphicspath{{../img/}}
+  \graphicspath{{img/}}
   \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
@@
 % *** SUBFIGURE PACKAGES ***
-%\ifCLASSOPTIONcompsoc
-%  \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
-%\else
-%  \usepackage[caption=false,font=footnotesize]{subfig}
-%\fi
+\ifCLASSOPTIONcompsoc
+  \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
+\else
+  \usepackage[caption=false,font=footnotesize]{subfig}
+\fi
 % subfig.sty, written by Steven Douglas Cochran, is the modern replacement
@@
 
+For techniques which involve applying a filter to decide the direction of a
+line within a picture, one often needs to be able to change or rotate the
+filter by varying amounts of degrees. This principal is often described as
+steering a filter and an efficient technique for doing so is described in 
+\cite{freeman91design} in both two and three dimensional space.
+
+This technique involves finding a function of $x$ and $y$ which steers a filter
+
+% TODO I'm struggling to understand this one...
+
 \subsection{Gabor Filters}
@@
 
+Another analysis paradigm is fractal geometry which is based on the idea that
+all analysis performed on an image at difference scales are equally important
+and that the richest information can be found in the mechanisms which relate
+them to one another.
+
+Because of this, fractal tools can be used to analyse contours and textures of
+images.
+
+Multifractal analysis concentrates on the use of processing tools which
+describe the fluctuations within regular regions of an object which at
+different scales.
+
+Until recently multifractal analysis was rarely applied to image processing
+problems, but with the breakthrough in an efficient formulation of
+multifractal analysis obtained through wavelet leaders.
+
+The use of wavelet leader multifractal analysis has been used to explore
+painting texture classification\cite{abry2013van}.
+
+Image processing is a field which has grown out of signal processing; an image
+can actually be thought of as a two-dimensional signal, which allows existing
+signal processing techniques to be applied to them. In early days of image
+processing the Fourier transform was often used to decompose images into
+signals. In more recent years wavelets and transforms using wavelets have
+become more popular, as they can provide all the features of a Fourier
+transform, but also provide localised time information, as well as localised
+frequency information.
+
+In the case of images, localised time information maps to the location within
+an image.
+
+In this research, a 2D discrete wavelet transform is applied to the image to
+gather the wavelets coefficients, normalising the image to the correct form
+needed for the definition of wavelet leaders.
+
+These coefficients enable a definition of the global regularity of the image.
+
+To allow image classification across different scales, one needs a dyadic
+space; a space in which there are a collection of regions of different scales,
+where a region at one scale can also be viewed as the union of regions in a
+smaller scale. Figure~\ref{fig:dyadic-image} shows an example of this where the
+large regions show only a pixalated view of the image and the smallest the
+individual pixels of the image.
+
+\begin{figure}[ht]
+\centering
+\subfloat[$C=8$]{\includegraphics[width=0.1\textwidth]{dyadic_2.png}}%
+\hspace{.2in}%
+\subfloat[$C=32$]{\includegraphics[width=0.1\textwidth]{dyadic_4.png}}%
+\hspace{.2in}%
+\subfloat[$C=128$]{\includegraphics[width=0.1\textwidth]{dyadic_8.png}}%
+\hspace{.2in}%
+\subfloat[$C=512$]{\includegraphics[width=0.1\textwidth]{dyadic_16.png}}%
+\caption{Example of dyadic space using the standard Lenna image (512x512
+pixels) at different dyadic scales ($C$), note how the picture becomes less
+pixalated at higher scales}
+\label{fig:dyadic-image}
+\end{figure}
+
 \subsection{Texton-Based Analysis}

### 2014-04-29 paper.tex
Add some more details.
Also changed to a less sexist picture. Hopper > Lenna.

@@
 where a region at one scale can also be viewed as the union of regions in a
-smaller scale. Figure~\ref{fig:dyadic-image} shows an example of this where the
-large regions show only a pixalated view of the image and the smallest the
-individual pixels of the image.
+smaller scale. Figure~\ref{fig:dyadic-image}\footnote{The photograph is
+property of the U.S. Federal Government, and is therefore in the public
+domain} shows an example of this where the large regions show only a pixalated
+view of the image and the smallest the individual pixels of the image.
 
@@
 \hspace{.2in}%
-\subfloat[$C=512$]{\includegraphics[width=0.1\textwidth]{dyadic_16.png}}%
-\caption{Example of dyadic space using the standard Lenna image (512x512
+\subfloat[$C=512$ (Original Size)]{\includegraphics[width=0.1\textwidth]{dyadic_16.png}}%
+\caption{Example of dyadic space using an image of Grace Hopper (512x512
 pixels) at different dyadic scales ($C$), note how the picture becomes less
@@
 \end{figure}
+
+A natural interpretation of multifractal analysis needs to be based on wavelet
+leaders, which allows an estimation of the multifractal spectrum of an image.
+The wavelet coefficients at finer scales within a small neighbourhood are used
+to renormalise the wavelet coefficients at a given scale.
+
+This is all theoretically very sound, but in practise it cannot be used on
+paintings without some modifications, mainly due to the fact that digital
+objects do not have an infinite resolution. So the analysis is slightly
+simplified to account for this to provide a fair estimation of the real
+analysis.
+
+This technique was applied to several different works, but the most interesting
+of there was the Princeton experiment, where an artist produced seven distinct
+small paintings using different materials. Two weeks later the artist was asked
+to produce replicas as close to the original as possible.
+
+Both the originals and replicas were scanned at a very high resolution to allow
+this analysis to be as detailed as possible.
+
+The results showed that, systematically, the textures of the replicas were
+globally more regular and smoother than the originals. However, it should be
+noted that getting these results required the expert selection of sections to
+analyse and the a posteriori selection of the range of scales for wavelet
+leaders.
+
+With promising results shown from this experimentation, attention was turned to
+a norm for painting analysis: Van Gogh. Rather than consider the global state
+of the canvas, each image was split into smaller sections for the analysis, as
+paintings very rarely have the same texture globally.
+
+This analysis was used to try and place paintings in a date region and detect
+forgeries. The results shown for dating paintings into a period show some
+promising results, where the majority of paintings were correctly clustered
+into their correct periods. For detecting forgeries the results are promising,
+but one forgery slips by when it was noticed by experts.
+
+This technique is agnostic to the colour space and the authors do apply it to
+several different colour spaces, including black-white intensity, \gls{RGB} and
+\gls{HSL}.
+
+There is a lot of expert knowledge which cannot be automated or, where it can
+be, is very arbitrary such as blindly selecting patches.
+
+The results from the Princeton experiments are promising, but the
 

### 2014-04-29 paper.tex
Finish paper 12.

@@
 
-The results from the Princeton experiments are promising, but the
+The results from the Princeton experiments are promising, but the results are
+very sensitive to the material of the canvas as well as the tools used, whilst
+this might be use for detecting forgeries, there might be situations where this
+isn't desired.
+
+The images the authors use are high resolution (800 DPI) and they note that a
+lower resolution makes it difficult to decide on the range of scales which are
+involved.
+
+The data sets used are relatively small and, although they increase the
+knowledge by taking multiple patches from the Van Gogh paintings, but for the
+results they show only individual paintings are considered. One could easily
+question the validity of their results based on this.
+
+In conclusion this is a very complex but powerful technique which is actually
+applying theoretical concepts to perform some interesting analysis and
+classification. However, the sensitivity of the technique and with the need for
+expert and a posteriori decisions do limit the real application of this
+research.
 

### 2014-04-29 paper.tex
Make a start on paper 11.

@@
 
+% TODO Again, another complex one :-(
+
 \subsection{Histograms of Edge Orientated Gradients}
@@
 
+Analysis of texture can also be performed with the help of a set of small
+patches, or textons, relating to the texture of the image. An interesting
+approach is to learn a variety of these textons from example images and then
+look at a histogram of the frequencies of these textons on the images to
+classify\cite{van2010texton}.
+
+Segmentation of brushstrokes is, understandably, difficult - especially
+digitally where the image is usually only two dimensional.
+
 \section{Statistical Analysis}

### 2014-04-30 paper.tex
Do paper 11.

@@
 \newacronym{HSV}{HSV}{Hue, Saturation, Value}
+\newacronym{PCA}{PCA}{Principal Component Analysis}
+\newacronym{t-SNE}{t-SNE}{t-Distributed Stochastic Neighbor Embedding}
 
@@
 Segmentation of brushstrokes is, understandably, difficult - especially
-digitally where the image is usually only two dimensional.
+digitally where the image is usually only two dimensional, but one can consider
+the texture of the painting with more ease and can provide a good insight into
+the artists style.
+
+Filters are becoming less popular for painting analysis as the filter often
+normalises the image to a certain extent in the process of analysis. Other
+approaches like wavelets and pixel-based representations, such as texton
+approaches, are becoming more popular to avoid this issue.
+
+It should be noted that whilst textons are a building block of the text of an
+image, they are not identical to brushstrokes.
+
+To construct the codebook of textons, 5000 patches are selected in random
+locations from each painting in the training set. These patches are then
+clustered and the most exemplary patch (the central patch) is selected as the
+most representative for that cluster.
+
+From this codebook it is then possible to generate a histogram for any image
+which estimates the distribution of these textons across the image. The
+histogram is created by applying a sliding window across the painting and
+selecting the nearest texton in euclidean space to the window. Finally the
+histogram is normalised to sum up to 1.
+
+The authors note that these histograms could be built up for different sizes of
+textons to increase the feature space available for analysis, for their own
+experiments they use six different scales of texton.
+
+Because the aim of this research was to work closely with experts in the field,
+the authors also came up with a method for visualising this information. For an
+image processing system this may not be required, but could be used to extract
+more useful information from the analysis.
+
+Because of the high dimension space it is necessary to reduce the
+dimensionality of the results to allow it to be human readable. Usually this
+would be performed by \gls{PCA}, but the way in which the data is structured,
+\gls{PCA} is too lossy for texton histograms; it has a linear nature whist the
+texton histograms may be high-dimensionally non-linear and it focuses on
+preserving the global structure rather than local structure of high-dimensional
+points.
+
+There are other techniques which might have been applied to solve these
+problems, but the authors found some shortcomings with these techniques which
+would have made real-life visualisation difficult.
+
+To cope with this, a new method for dimensionality reduction was invented:
+\gls{t-SNE}. This method hinges around keeping
+the conditional probabilities between the high and low dimension spaces
+similar.
+
+Interestingly, in the experiments \gls{PCA} was first use to reduce the
+dimensions down to 50, and then \gls{t-SNE} was
+used to reduce down to two dimensions. Suggesting that, for very high dimensions,
+\gls{t-SNE} is not very effective.
+
+The results on 117 high-resolution grayscale Van Gogh paintings show that most
+of the non-Van Gogh paintings appear on the peripheries of the visualised
+two-dimensional graph, apart from two specific forgeries: the Wacker forgery
+which had fooled experts for many years, but can successfully be found by
+looking at global features. The other, created by Gaugin, may remain undetected
+for the same reason.
+
+Visualisation of dated Van Gogh work does show some difference between the two
+time periods considered, but not enough to cluster or classify upon with any
+degree of accuracy.
+
+% How many clusters?
+
+This approach is definitely a useful one, especially given that the textons are
+learned from existing work, making analyses such as trying to date an artists
+work from sets of his known work very applicable. Although colour and global
+features are not considered by the textons described in this work, colour would
+be a simplistic feature to add and global features could be part of a separate
+technique which is later included with the histogram.
+
+The paper does leave some answered question, one notable one is how best to
+decide the number of clusters to build the texton histogram from and which
+scales the textons work best at. For use by experts with \gls{t-SNE} applied
+this might be some useful work, but \gls{t-SNE} seem to have limited
+application outside this form of data set. A comparison between \gls{PCA} and
+\gls{t-SNE} would, perhaps, show the advantage of the latter and an explanation
+as to why \gls{PCA} was applied before \gls{t-SNE} on their own experimental
+results would help others to better apply such methods correctly.
 

### 2014-04-30 paper.tex
Blah...

@@
 
+A lot of analysis can be gained from passing filters over an image and
+considering the colours used. However, a lot of these depend on some way of
+statistically analysing the results. This section will look at some of the
+research that considers some of the more statistical elements to provide
+analysis.
+
+Whilst gaining results of analysis techniques, it is also important to be able
+to interpret them digitally for some form of use; determining whether a
+painting is authentic or not, or dating an artist's work within a set of their
+known work.
+
 \subsection{Stylistic Analysis}

### 2014-05-01 paper.tex
Add paper 9 details.

@@
 
+Often these techniques involve using a form of machine learning to gather
+meaningful data from the high dimensional feature space which image processing
+processes provide.
+
 \subsection{Stylistic Analysis}
 % Paper 9
+
+All artists have a distinguishing style which may change as their career
+progresses. Art historians find this a challenging problem as many factors of
+the painting need to be taken into account. Image processing and machine
+learning techniques can be used to aid this process.
+
+Stylometry, the study of an artists style, has many problems involved with it.
+Two of these are: extracting distinguishing features and dating a
+painting, both these problems are tackled in \cite{jafarpour2009stylistic}.
+
+This research uses considers the artists style to be a hidden variable which
+controls the observable properties of the image; the colour, brushstrokes, etc.
+
+The dating challenge is the act of dating a painting from a catalogue of known
+paintings from the artists lifetime. An artists style is likely to change over
+time, especially as they meet their peers and try out new movements.
+
+This research considers how a human expert might approach the problem of
+determining style and that paintings may degrade over time due to the materials
+used to create them. Historians must combine a potentially noisy observation
+with pre-existing knowledge to come up with an analysis of the style.
+
+Computer systems have to use features of an image, considering both global and
+local information, to build up a high-dimensional feature space which can then
+have statistical analysis performed upon it to gauge some information about
+style.
+
+It has already been discussed that a \gls{HSL} colour space provides a more
+representative view of an image than a \gls{RGB} space. However, \gls{HSL} does
+not exist in a Cartesian space. This can be achieved by transforming the
+\gls{HSL} space into an XYZ space by ``unrolling'' the radial elements of
+\gls{HSL}.
+
+As with many techniques which utilise wavelet transforms, a very high
+resolution is used to gather information which may be too fine for a human eye
+to perceive. A dual-tree complex wavelet transform using the aforementioned XYZ
+colour space can detect colour patterns as well as local difference.
+
+These wavelets provide coefficients in very high dimensional space which a
+large amount of noise. To deal with this there is a need for dimensionality
+reduction and normalisation of noise within the feature space.
+
+Hidden Markov Trees provide an image model over numerous different resolutions.
+They act like Hidden Markov Models but also considers a tree like structure
+which maps to dyadic space. This allows for an image to be described as a
+statistical entity rather than a set of pixels.
+
+At each scale there are hidden variables which control the wavelet
+coefficients. These hidden variable represent either a smooth region, which has
+a small variance, or an edge, which has a large variance.
+
+The images are split into several patches to allow the analysis to be performed
+effectively.
+
+Results for the dating challenge, 66 Van Gogh paintings were used with the goal
+of dating three test images which cannot be easily dated by art historians.
+Training was performed on the set of paintings using 10-fold cross-validation
+using several different classifiers. The best classifier; random forest; had a
+generalisation performance of 73.7\% and the results from the three test images
+aligned with the conclusions of the art historians, but this is somewhat
+conjecture as the dates aren't officially known.
+
+Extraction distinguishing features focused on extracting flowers within the
+paintings. Irrelevant patches were removed from the training images and again
+training was performed using 10-fold cross-validation using several
+classifiers. Once again random forest was the best decision and several
+distinct distinguishing features.
+
+The results of this research do seem promising, and the techniques described
+here do seem like they can provide some very powerful analysis.
+
+The sample size is a little small, but the split into different patches allows
+for a larger number of samples for training. Though this still only has a
+certain number of date ranges to work with.
+
+The size of the patches is never fully explored and the research blindly uses
+$256 \times 256$ pixels for each region. It would be interesting to see what
+effect changing the size of these regions would make upon the research.
+
+In conclusion this research provides some very powerful analysis with the
+experimental results to back it up in two different fields of stylometry which
+agree with expert opinion and begin to bring forth some new ideas.
 

### 2014-05-01 paper.tex
Add paper 4.

@@
 \newacronym{t-SNE}{t-SNE}{t-Distributed Stochastic Neighbor Embedding}
+\newacronym{MDS}{MDS}{Multidimensional Scaling}
 
@@
 
+Another interesting problem within the field of art analysis is the problem of
+many hands; that is determining how many authors created a work and which parts
+of said work can be attributed to which author. Obviously, the latter can also
+be simplified to test which artist created a work and thus detecting forgeries.
+
+One attempt to attempt to solve the many hands problem is described in
+\cite{lyu2004digital}. As with much of the recent work into applying image
+processing to artwork, it considers paintings at multiple resolutions and
+orientations using wavelets.
+
+% TODO technique.
+The method for this technique is to decompose the image using wavelets. In this
+paper quadrature mirror filters; filters which split an input signal into two
+bands; are used to split the frequency space into multiple scales and
+orientations. This produces a statistical model at each scale and orientation;
+known as a subband.
+
+The optimally predictive set of neighbours is selecting by minimising the
+prediction error using a brute force iterative approach on a per-subband and
+per-image basis across all subbands.
+
+Additional statistics are also gathered from the errors of the final predictor,
+again for each subband. For $n$ levels of decomposition, this will produces
+$24(n-2)$ values ($12(n-2)$ for optimal predictors, $12(n-2)$ for error
+statistics).
+
+These large dimensional spaces can be projected into a human-friendly subspace
+using a process named \gls{MDS}. \Gls{MDS} attempts to maintain the distance
+between data points when projecting into a lower dimensional space so should
+still provide a decent visualisation of the higher dimensional space.
+
+The first experiment focused on authenticating the work of Pieter Brugel the
+Elder who used very distinct style which the authors felt would be especially
+applicable to their technique. A set of 13 paintings conprising of 8 authentic
+Brugels and 5 famous imitations was the subject for this analysis.
+
+Each painting was scanned in very high resolution and was separated into 64
+non-overlapping regions. Each subimage was transformed the proposed wavelet
+transform at five levels and three orientations producing a 72 dimensional
+space. Authentication was performed by applying Hausdorff distance to this
+space with the assumption that authentic paintings would be closer together.
+
+This these experimental results it did indeed show that this was the case and
+are statistically significant, even in a reduced feature space through
+dimensionality reduction.
+
+This was also to applied to a painting by Pietro di Cristoforo Vannucci
+(Perugino) which is disputed by art historians to have been produced by more
+than a single artists, which, during the Renaissance was common amoungst the
+great works.
+
+This effect is commonly observed in the six faces which are depicted in this
+painting. These faces were first selected and then partitioned into $256 \times
+256$ regions as with the Brugel work. The technique was then applied and
+Hausdorff distance was measured between the feature spaces. Viewing the results
+shows an obvious clustering of three of the faces, whilst the other three are
+completely separate from this cluster and from each other. This would suggest
+that there were at least four hands involved in the painting, which is
+consistent of the views of some art historians.
+
+The sample size both experiments were performed on is relatively tiny when
+compared to other studies using similar techniques. One could easily question
+the validity of such results, especially with the subjectivity of the second
+experiment where there is no definitive answer.
+
+The work of Brugel is noted to have delicate lines and shading which was
+expected and shown to have good results. A cynical reader might infer that this
+analysis technique is therefore not suited to other styles of art.
+
+It's also interesting that the Brugel paintings were cropped to a given size
+before being split into regions as this would surely lead to loss of
+information when it might have been better to just increase the number of
+regions across the image to include the whole image.
+
+As with other research including wavelets the dependence on a region size of
+$256 \times 256$ pixels is notable. This appears to be a standardised size for
+the application of wavelets, but there appears to be no documentation of this.
+
+In conclusion this work presents a method which may provide an insight into how
+many authors a work has, but has very few experimental results to back up this
+claim. Especially given other research into similar areas.
+
+% Line shaded -> applicable to paint?
+% Crop size -> loss of information
+
 \subsection{Dating an Artist's Work}
@@
 % Paper 2
+
+\section{Conclusions and Summary}
+% TODO beast mode this.
 

### 2014-05-03 paper.tex
Add introduction for brushstrokes.

@@
 
-\section{Texture Analysis}
+\section{Texture Analysis}\label{sec:texture}
 Paintings are somewhat unlike the normal subject for image processing, whilst
@@
 
+The ideal technique for analysing a painting digitally is to be able to extract
+brushstrokes from the painting and be able to analyse them based on their
+shape, colour and where on the painting different types of stroke were made.
+
+This is a very difficult task, it may be easy for a human non-expert to view a
+painting and roughly estimate where the surface brushstrokes are, implying that
+an expert may be able to tell far more from a detailed view of the painting,
+including multispectral imaging using X-rays.
+
+For a computer to do this is a very intensive and complex task. As discussed
+in section~\ref{sec:texture} is can be easy to view the texture of an image and
+even to be able to pull out some information about the texture of brushstrokes.
+But actually segmenting these brushstrokes for extraction quickly is a
+different story.
+
+This section will look into some of the research which has been attempting to
+extract brushstrokes from paintings for the purposes of identification and
+authentication of artwork. However, because this field is relatively narrow,
+some exploration into extraction of brushstrokes in calligraphy will also be
+considered as the fields have some overlap.
+
 \subsection{Artistic Identification}

### 2014-05-03 paper.tex
Add paper 1.

@@
 
+\usepackage{paralist}
+
 \usepackage{url}
@@
 \newacronym{MDS}{MDS}{Multidimensional Scaling}
+\newacronym{SVM}{SVM}{Support Vector Machine}
 
@@
 
+To be able to extract brushstrokes it is important to understand how an expert
+might go about the task. A lot of analysis techniques tend to be inspired from
+how a human might go about the task, this also help a human to understand how
+the process was gone about and may give extra meaning to the results.
+
+There is a lot of knowledge to be gained from looking at the expert process,
+and this knowledge can be applied to reduce noise and parameters from the
+extraction process. The knowledge of painting materials, the state of the
+preservation of the painting and the working methods of the artist themselves.
+
+Interestingly, this leads to the conclusion that parts of the painting should
+actually be discounted from the analysis, elements which have deteriorated over
+time due to instabilities in the painting materials or the lack of preservation
+of the painting leads to what most computer vision researchers would call
+``noise''. Certain areas of the painting may not have been painting by the
+artist's hand, such as cracks in the painting from the drying process, need to
+be considered here.
+
+It seems that it should be far more important to look at the areas where
+brushstrokes were obviously painting ``rhythmically'': a series of fast strokes
+arranged in a certain pattern, which can then be used to build up a personal
+handwriting for a given artist\cite{johnson08image}.
+
+Three different two-dimesnional wavelet transformation techniques were used in
+this research:
+
+\begin{itemize}
+\item Penn State
+\item Princeton
+\item Maatrich
+\end{itemize}
+
+Penn State is an orthonormal series generated by a wavelet, implemented using
+fast subband filtering, critically sampling the data.
+
+Princeton forgoes critical sampling in favour of greater orientations using
+complex wavelets.
+
+Maatrich uses Gabor wavelets to provide a similar range of orientation
+selectivity as Princeton, but in a way which is closer to a physiological
+templates.
+
+Because of the differences between these wavelet transforms, different
+classifiers need to be applied to the resulting coefficients.
+
+Both Penn State and Princeton transforms can use a hidden Markov model
+approach. Because Markov models have states which are linked to the predecessor
+or neighbour states with a probability distribution. Hidden Markov models are
+Markov models where this probability distribution is unknown and can be
+inferred from the data which constructs the model.
+
+There are several classifiers which can be applied to Markov models, although
+typically these relate to some form of expectation maximisation technique.
+
+All experimental results were gathered from a data set containing 101 paintings
+in the style of Van Gogh. 82 of which are believed to be true Van Gogh works,
+13 which are currently in dispute and 6 which are known imitations. All of
+these paintings are scaled using the same technique to be equivalent sizes.
+
+The approach for using Penn State was to gather several unquestionably Van Gogh
+paintings which are representative of his style (exemplar images) as a training
+set. By not including negative examples in the training set the resulting
+classifier is more stringent.
+
+The paintings are then divided into patches of roughly $512 \times 512$ pixels
+such that all the patches from the same painting are the same size.
+Dissimilarity measures between these patches are calculated using both texture-
+and stroke-based features. Paintings are then compared using the aggregated
+distances between the patches of the images.
+
+Texture features were extracted in a form which shows the abruptness of changes
+in variations at different orientations. Textures were only captured at the
+highest scale. These are then put into a feature vector which can be converted
+into a Markov model.
+
+For extraction of strokes an edge-detection algorithm is used to detect
+contours of the image, which can then have several geometric features computed,
+such length, average curvature and orientation.
+
+Because textures are dependent on their neighbours it is sensible to build up a
+probabilistic model, namely a Markov model, to show these dependencies. Because
+the probabilities which affect the dependencies are unknown this must therefore
+be a hidden Markov model, using two dimension (as images are two-dimensional).
+A likelihood classifier can then be trained from the training set and applied
+to other 2D hidden Markov models to be able to classify examples.
+
+Spatial features are less important for brushstrokes as they capture more
+global information rather than information which is likely to be affected by
+neighbouring strokes. Clustering can be applied to this to partition the data
+set then new examples can be classified using the distances to these clusters.
+
+Princeton shows the scale at which details start to emerge; each patch is
+analysed using the wavelet transform and then these coefficients are stored is
+a hidden Markov tree in a similar process to \cite{jafarpour2009stylistic},
+where the hidden states take the form of edge or non-edge. This helps show the
+scale at which an individual artist's style emerges as the transitions between
+edge and non-edge will occur at different scales.
+
+Only 79 of all 101 paintings were analysed by Princeton, the other paintings
+were too dark for meaningful analysis to be gained from them in this
+experiment. 65 of these were van Goghs, 6 imitations and 5 of the questioned
+paintings.
+
+\gls{MDS} is used to find an arrangement of all the points in a three
+dimensional space (as opposed to the very high dimensional space the hidden
+Markov tree uses. For all van Gogh paintings the centre point is determined,
+where all of these paintings are equally weighted.
+
+The 3D visualisation shows a good separation of the van Goghs from the non-van
+Goghs, but to be able to truly use this, a radius classifier is applied such
+that a lower radius is more likely to classify as a van Gogh and visa versa. An
+appropriate point $r$ is chosen as the cut distance at which the label changes,
+such at van Gogh paintings are closer to the centre than $r$.
+
+This classifier was tested using leave-one-out cross validation; a validation
+technique based on $k$-fold cross validation where $k$ is the size of the data
+set. $k$-fold cross validation partitions the data set into $k$ subsets. Each
+subset has classification performed on its elements based on training from the
+other subsets.
+
+Leave-one-out cross validation is a useful cross-validation technique for small
+data sets and can be used to gather a correlation of expected output against
+classified result. This can also be used to gather the statistical significance
+of the results.
+
+Biologically inspired techniques have shown promising in other fields of
+computer science, notably items like Artificial Neural Networks, Immune and
+Endocrine Systems; Swarm Intelligence such as Ant Colony Optimisation; etc.
+
+Using the human visual system as a metaphor paintings can be analysed using
+three principals: 
+\begin{inparaenum}[1)]
+\item Importance of contours;
+\item Analysis at multiple scales;
+\item Similarities are reflected in local textures.
+\end{inparaenum}
+
+The Maatrich transform uses Gabor wavelet filters at multiple scales and
+orientations. The provides an ``energy'' value for each pixel, orientation and
+scale. These values can then be binned into a high-dimensional histogram. This
+allows a \gls{SVM} to be trained and new example to be classified
+with. Because there are only two states (van Gogh and non-van Gogh), a
+\gls{SVM} is a good classifier. Again, leave-one-out cross-validation is used
+for each painting, the patches are separately classified then the most
+frequently returned label is used to classify the painting.
+
+% PS results
+Penn State begins to show some good results, although one of the imitations is
+considered to be one of the five closest to the training set of 23 van Goghs.
+
+% Pr results
+The radius classifier applied to the Princeton tranformation successfully
+classified 55 of 65 van Goghs and 9 out of 11 non-van Goghs correctly. At even
+finer scale shows even better results.
+
+% MA results
+Maatrich transforms with a \gls{SVM} classifier successfully classified four of
+the six non-van Goghs correctly, but wrongly classifying two of the van Goghs.
+This starts to show that it can detect dissimilarities in the brushstroke
+texture of a painting, but that some difference may be too subtle to pick up.
+
+% Conclusions
+The use of leave-one-out cross-validation means the experimental results are
+very valid; but the lack of significance does remove some of this validity - it
+is up for debate whether a null hypothesis could also be to play.
+
+The images are high resolution greyscale, so although colour information has
+been lost but the resolution means that most of the paintings should be uniform
+and detailed.
+
+The discussion and use of expert knowledge means that the results gained are
+potentially more valid by not considering several sections of a painting.
+However, the opposite argument could be made that part of van Goghs style was
+to deform the paintings slight by leaning against them, etc. and therefore is a
+better indicator of his style than some of the brushstrokes.
+
+This does show positive results and the techniques are well thought out,
+including discussions of other classifiers which could have been used.
+
+As is a common problem in brushstroke extraction there is no ground truth on
+which these techniques can be applied to, so it is difficult to know how well
+these techniques will apply to other paintings by van Gogh or by different
+artists.
+
+In conclusion, this is a decent application of existing research with real
+results being gained, although it doesn't deal with the extraction of
+brushstrokes formally it is a good step in the right direction.
+
 \subsection{Rhythmic Brushstrokes}
 % Paper 2
+
+\subsection{Stroke Decomposition}
+% Paper 15
+
+\subsection{Calligraphy Stroke Extraction}
+% Paper 14
 

### 2014-05-07 paper.tex
More changes :-)

@@
 
-% TODO technique.
 The method for this technique is to decompose the image using wavelets. In this
@@
 
+Using computer vision to segment brushstrokes is a difficult task; often
+requiring expert knowledge of the painter and manual intervention to enable the
+process. \cite{jia12rhythmic} shows a method for the automatic extraction of
+brushstrokes.
+
+This method involves a complex process; first the image has an edge detection
+algorithm applied to it, then these edges are pruned to remove noisy, short and
+unconnected edges to improve the actual segmentation of strokes.
+
+Because an edge detector doesn't always pick up the full brushstroke, any gaps
+in the painting are enclosed using a greedy process of scanning neighbours and
+linking the nearest via a straight line if they are within a certain distance.
+
+Any fully enclosed brushstrokes are then extracted and checked against an upper
+and lower threshold for size. These thresholds were pre-selected to give a good
+estimate for the size of a brushstroke. This is finally labelled a brushstroke
+if the backbone is not too branched, it has a reasonable ratio of broadness to
+length and the size of the stroke is proportional to the backbone of the
+stroke.
+
+After this process is complete, the image is then segmented with regard to
+colour (\gls{RGB}) and directions (horizontal and vertical). $k$-means
+clustering is then applied multiple times, gradually decreasing the threshold
+for the average distance in the cluster. Any sufficiently small clusters are
+removed at each step and are not considered in the future runs of the $k$-means
+algorithm. This process is done to remove any noisy components there might be
+in the digitisation of the image.
+
+Brushstrokes which are already labelled are then removed and any remaining
+clusters are labelled according to the same process.
+
 \subsection{Stroke Decomposition}

### 2014-05-07 paper.tex
Add paper 2.

@@
 
+Calculating the backbone is based on the edge detection of a segmented
+brushstroke, this will generally form ``Y'' shaped structures; in the case of a
+few of these structures, the shortest branch is simply removed. This process
+might produce a disconnected backbone, at which point it is labelled
+``severally branched''.
+
+To gain experimental results, 45 digitised oil paintings from van Gogh and his
+contemporaries were used, ranging in sizes and scales. A ground truth for the
+brushstroke extraction was generated manually from 10 regions of these
+paintings.
+
+First, the system used to decide if a region is a brushstroke was evaluated on
+these manually generated brushstrokes, obtaining a correct decision 95\% of the
+time.
+
+Then, the automatically extracted brushstrokes were compared to the manually
+selected ones in terms of: validity (the percentage of extracted brushstrokes
+which are in the ground truth) and detection (the percentage of the ground
+truth strokes which were extracted).
+
+Validity reached an average of 60\%, but ranged between 42\% and 83\% for
+different paintings. Detection also reached and average of 60\%, but the lowest
+was 22\% and only reached a high of 75\%.
+
+Of course, the extracted brushstrokes also need analysis performed on them.
+This is where the backbone calculation can also be used to extract features out
+of the brushstroke, as well as considering the features of the brushstroke
+itself.
+
+Using this statistically, this can be used to authenticate paintings and also
+begin to place them in different periods. However, the results from these are
+in terms of $p$-values for each feature and do not consider multiple features.
+However, it was found that the features which distinguish van Gogh from his
+contemporaries are not the same as the ones which differentiate two different
+periods in van Gogh's career.
+
+These results are very promising, especially given that the brushstrokes were
+compared to a ground truth. Of course it could be argued that the ten regions
+selected were done so to show the best of the technique, but given the
+visualisation of the brushstrokes, this seems unlikely.
+
+It would have been interesting research to see how the threshold values for the
+detection of a brushstroke; there is no investigation or discussion as to why
+the high and low values were set where they were.
+
+The same goes for $k$-means clustering; is there a better choice which provides
+more accurate clusters, and therefore strokes, or was it chosen because of its
+simplicity and the benefits gained from using a more complex one wouldn't be
+worth the computational complexity?
+
+One would have thought that a density-based clustering algorithm such as
+DBSCAN\cite{kriegel11density} would have been more applicable to this problem
+as it would have been able to cope with noise and may have provided a better
+segment of the brushstrokes.
+
+There is also the question of applicability. That is, this has been 
+successfully applied to van Gogh and some of his contemporaries with oil
+paintings, but would the technique still yield such good results is applied to
+an artist who used a palette knife, etc.?
+
+In conclusion, this paper presents a very useful method for extracting
+brushstrokes and compares the process of doing so to a ground truth. There are
+still some elements which could use further investigation, but on the whole
+this is a very rounded piece of research.
+
 \subsection{Stroke Decomposition}

### 2014-05-07 paper.tex
Start on paper 15.

@@
 
+One interesting use of stroke extraction is to animate Chinese paintings from
+an original\cite{xu2006animating}. This is a complex procedure which involves
+decomposing the painting based on brushstrokes, these brushstrokes can be
+represented as vectorised elements and become easy to manipulate and therefore
+animate.
+
+To decompose the strokes the following method is proposed.
+
+First, the image is over-segmented using a mean-shift algorithm, this is done
+because colour is required to extract brushstrokes, but is not enough alone.
+The over-segmentation stops regions spanning across multiple strokes, unless
+the strokes themselves are overlapping.
+
+The strokes are then extracted by merging these regions until they represent
+the brushstroke. This requires some domain knowledge as naively merging these
+regions may not represent the stroke accurately enough.
+
+Two intuitions are used in this research: colour gradients are low along brush
+strokes and strokes will match a library of strokes manually extracted from
+other sources.
+
 \subsection{Calligraphy Stroke Extraction}

### 2014-05-07 paper.tex
Add paper 15.

@@
 
+To merge the segmented regions a region merging criterion $\epsilon$ is
+calculated. $\epsilon$ is then used to merge adjacent regions from highest to
+lowest. $\epsilon$ is calculated in terms of the differences in colour
+distribution between the regions and the similarity to the catalogue of
+brushstrokes.
+
+Once this process is complete, the strokes can then be refined and vectorised
+for the animation process.
+
+This paper also considers the separation of overlapping strokes to be able to
+animate them naturally, but this could potentially be used for further
+analysis. The overlapping regions can be identified in the segmentation process
+then have an iterative algorithm which separates the colours of the overlapping
+region. Because the colouring of an overlapping region can be the mixture of
+more than a single colour, a mixture model is used to model the overlapping
+strokes.
+
+Combining this with the library of known strokes can be used to get a good
+estimate of the separate strokes.
+
+This paper presents some really interesting research into the decomposition of
+paintings using brushstroke extraction; but is very specific to Chinese
+paintings. An example failure run on a painting by Georges Seurat shows the
+overfitting which this method falls prey to.
+
+The reliance of a catalogue of brushstrokes seems to do more harm than good in
+some circumstances, the authors do note that the catalogue is by no means
+exhaustive. This is what seems to lead to the over segmentation. Perhaps
+learning the strokes from existing work as discussed in \cite{van2010texton}
+could provide more generalised results.
+
+Unlike \cite{jia12rhythmic}, the extracted strokes are not evaluated against a
+ground truth. Even so, the visualisation of the segmented brush strokes seems
+reasonable and to be able to complete the animation steps there must be some
+validity to the extraction process.
+
+The development of a technique to discriminate two overlapping strokes shows
+some potential, although there is no real verification of the method to do
+some. It should also be noted that the paper only considers two overlapping
+strokes as more would become too computationally complex to complete in a
+reasonable time.
+
 \subsection{Calligraphy Stroke Extraction}

### 2014-05-07 paper.tex
Add paper 14

@@
 
+Chinese calligraphy is a painted art in its own right, although not exactly
+like a painted scene it does provide some insight into stroke extraction which
+has a much more formalised ground truth.
+
+\cite{xu2005automatic} described a method for the automatic extraction of
+calligraphy strokes, with account for differing stroke sizes, deformations and
+noise found in historical work.
+
+Unlike paintings, Chinese calligraphy is only comprised of a certain catalogue
+of strokes and are categorised into either basic or composed strokes.
+
+The input, a binary image, has the Canny edge detection
+algorithm\cite{canny86computational} performed on it to extract the edge of the
+character.
+
+Because calligraphy strokes often overlap, the corner points of the edges of
+the character are found preserving information about the direction this process
+was performed in.
+
+A half-edge colouring graph can be created to show the relations between
+different edges. Each edge is split into two half-edges which has colours
+according to the relation to other edges. This information, plus the direction
+information allows the computation of matching different edges to build a
+single stroke, with relation to the distances of different vertices.
+
+This is done by removing the largest edge then trying to find one or more edges
+which can match to it to make a stroke. Once the stroke is found it can then
+have the gaps filled in by linking the two nearest vertices.
+
+There is little experimentation and no formal validation performed other than a
+visual check. The authors even note an occurrence where the corner detection
+algorithm failed to provide reasonable vertices and so the strokes had to be
+extracted manually.
+
+One would have though in an area where it would be easy to generate a ground
+truth, it would have been used to at least perform some form of statistical
+analysis on the results gained.
+
+This might be useful in edge cases of digitally analysing paintings, but
+without knowing how well it truly performs on calligraphy, one might be
+hesitant to apply this to a what is a much more complex field which needs the
+consideration of colour as opposed to a binary intensity.
+
+Whilst there are claims this technique should be able to cope with noise from
+historical degradation of the image, it is difficult to see how. Especially as
+Canny is known to susceptible to noise and can be improved to be less
+sensitive\cite{xiao10improved}.
+
 \section{Conclusions and Summary}

### 2014-05-07 paper.tex
Van Gogh -> van Gogh.

@@
 In the experimental results this paper provides, the latter is achieved by a
-study of another of Van Gogh's works from the same time.
+study of another of van Gogh's works from the same time.
 
@@
 With promising results shown from this experimentation, attention was turned to
-a norm for painting analysis: Van Gogh. Rather than consider the global state
+a norm for painting analysis: van Gogh. Rather than consider the global state
 of the canvas, each image was split into smaller sections for the analysis, as
@@
 The data sets used are relatively small and, although they increase the
-knowledge by taking multiple patches from the Van Gogh paintings, but for the
+knowledge by taking multiple patches from the van Gogh paintings, but for the
 results they show only individual paintings are considered. One could easily
@@
 
-The results on 117 high-resolution grayscale Van Gogh paintings show that most
-of the non-Van Gogh paintings appear on the peripheries of the visualised
+The results on 117 high-resolution grayscale van Gogh paintings show that most
+of the non-van Gogh paintings appear on the peripheries of the visualised
 two-dimensional graph, apart from two specific forgeries: the Wacker forgery
@@
 
-Visualisation of dated Van Gogh work does show some difference between the two
+Visualisation of dated van Gogh work does show some difference between the two
 time periods considered, but not enough to cluster or classify upon with any
@@
 
-Results for the dating challenge, 66 Van Gogh paintings were used with the goal
+Results for the dating challenge, 66 van Gogh paintings were used with the goal
 of dating three test images which cannot be easily dated by art historians.
@@
 
-This was also to applied to a painting by Pietro di Cristoforo Vannucci
+This was also to applied to a painting by Pietro di Cristoforo vannucci
 (Perugino) which is disputed by art historians to have been produced by more
@@
 All experimental results were gathered from a data set containing 101 paintings
-in the style of Van Gogh. 82 of which are believed to be true Van Gogh works,
+in the style of van Gogh. 82 of which are believed to be true van Gogh works,
 13 which are currently in dispute and 6 which are known imitations. All of
@@
 
-The approach for using Penn State was to gather several unquestionably Van Gogh
+The approach for using Penn State was to gather several unquestionably van Gogh
 paintings which are representative of his style (exemplar images) as a training

### 2014-05-09 paper.tex
Just conclusion to go :-)

@@
 
-This technique involves finding a function of $x$ and $y$ which steers a filter
-
-% TODO I'm struggling to understand this one...
+This technique involves finding a function of $x$ and $y$ which rotates a given
+filter by a given angle, typically in terms of $\frac{\pi}{4}$ rotations, but
+it can be more fine-grain than that.
+
+A lot of this research stems from the idea that images are really
+two-dimensional signals; therefore filters can also be thought of similarly.
+And a function to rotate them will generally deal with shifting the signal
+along an axis or transforming the signal so it has the effect of rotating the
+signal.
+
+This process can have high computational costs when used on a filter which is
+not separable in terms of $x$ and $y$, so it is preferred if a filter can be
+designed to be so, rather than take the computational cost.
+
+These filters can be used for several applications, from analysing the local
+orientation of gradients, which in turn can be used to build up systems which
+detect items in images, such as \cite{dalal05histograms}; help to remove noise
+in an image based on angles; detect contours within an image when applied to an
+image transformed using an edge detection algorithm; and calculating the shape
+from the shading of an image.
+
+This paper presents an idea and its application but reads more like a textbook
+reference than an experiment. As such there are no actual results detailed by
+the paper other than to visualise some of the problems and solutions steerable
+filters can be applied to.
+
+Of course one could question the applicability of this research; but given the
+wide-spread use of this research\footnote{Cited by at least 680 other IEEE
+papers} this is very doubtable.
 
@@
 
-% TODO Again, another complex one :-(
+Gabor filters are a sinusoidal wave within a Gaussian
+envelope\cite{gabor1946theory}\cite{daugman1985uncertainty} which allows the
+filter which matches signals with variable orientation, frequency and
+bandwidth. They are useful within image processing as they can be steered in a
+near limitless range of orientations by the change of a single parameter. The
+size (frequency) of the areas can also be tuned with the modification of a
+second parameter, making them very powerful.
+
+These attributes make Gabor filters appropriate for the task of texture
+discrimination. Texture discrimination is the task of labelling whether two
+textures are the same or different. To the human vision model this is quite a
+natural and simple take, but to a machine this is a complex task as a simple
+equality check does not account for noise or changes in intensity.
+
+Gabor filters are designed to model the way human vision system works as
+closely as possible.
+
+\cite{turner1986texture} describes a method of using Gabor filters for texture
+discrimination and is done by passing Gabor filters at $\frac{\pi}{4}$
+increments in orientation from $0$ to $\pi$ over overlapping regions of images
+at two phases. The results of the passes of each filter are then made into a
+phase insensitive measurement by taking the Euclidean distance of the two
+measurements based on the result from the filter and the intensity of the
+pixels.
+
+Regions are then partitioned based on the sum of this measurement across a
+square of pixels.
+
+Although the results are very subjective; based on visualisations compared to
+the human analysis, they do show some promising results, although in some
+respects not entirely unexpected.
+
+For example, Gabor filters ``see'' T-shaped structures closer to L-shaped
+structures rather than T-shaped structures rotated by $\frac{\pi}{4}$. This
+could be due to the low range of orientations the authors use in this paper.
+
+This paper does show how useful Gabor filters can be in their application to
+image processing; especially when considering the texture of images; however it
+does also expose some of the flaws of a filter-only approach.
+
+Although only using four orientations may appear to limit the applicability of
+the filters, other research\cite{brown13can} also shows that higher
+orientations does not always provide better results.
 
@@
 
+Another avenue which can be taken is to use an artist's existing work to try
+and classify new examples. \cite{brown13can} describes one attempt to do this
+using a variety of techniques, from colour analysis to texture details.
+
+To verify the results gained from this research the authors used a
+leave-one-out cross-validation strategy on the 102 works for which a date was
+know. This provided a correlation between actual year and classified year, as
+well as the statistical significance of these results and the number of
+paintings which were successfully classified within 15 years.
+
+Analysis techniques provided results in the form of histograms which could then
+be used to measure the distance between an example and the training set and
+classified using $k$-nearest neighbour. This classification algorithm was used
+because of its simplicity and naivety, which allowed it to only consider
+distance in the feature space rather than any patterns in the underlying data.
+
+Graphing the results of the classification a ``best'' value of $k = 7$ for
+$k$-nearest neighbour was selected visually and used to gather formal results.
+These results show correlations approaching $0.5$ for some of the more advanced
+techniques; those examining the texture of the work.
+
+This research also considered the idea of exemplar paintings; those paintings
+which are most representative for a given year. Expert knowledge was used to
+produce a list of \emph{Artistic Exemplars}, those most representative of the
+artists style to an expert in the field. From analysis of a group of paintings
+for each year a central point was computed, then the closest painting to this
+central point was chosen as a \emph{Statistical Exemplar} for that year.
+
+These exemplars were then used to classify other paintings by labelling the
+painting with the nearest exemplar. Though these don't show as good correlation
+results, it does allow other avenues of work to be created such as
+investigating the differences between the \emph{Artistic} and
+\emph{Statistical} exemplars.
+
+This work does start tcentral point was computed, then the closest painting to
+this central point was chosen as a \emph{Statistical Exemplar} for that year.
+
+These exemplars were then used to classify other paintings by labelling the
+painting with the nearest exemplar. Though these don't show as good correlation
+results, it does allow other avenues of work to be created such as
+investigating the differences between the \emph{Artistic} and
+\emph{Statistical} exemplars.
+
+This work is obviously in its infancy but does show some promising results. The
+methodology is well thought through.
+
+The authors do discuss the use of $k$-nearest neighbour and whether it should
+be compared to other classifiers. An \gls{SVM} seems particularly applicable as
+it would help draw out some of the underlying patterns in the artists work.
+
+The lack of high-resolution images with non-standard lighting has hindered this
+research somewhat; although part of the point of the research is that the
+system should be robust enough to handle simple catalogue images, it does
+appear to damage the performance.
+
+It is also very specific to the artist Kyffin Williams, as his work was highly
+textural, which might explain why features including texture performed better.
+More work could be done to explore other artists and see how they compare to
+Williams.
+
+As the work uses histograms, it might have provided better results to combine
+the results of the analysis techniques to gather a larger histogram. This would
+have to be normalised or weighted based on the performance of the techniques to
+gather meaningful analysis, but is another avenue which hasn't been explored
+sufficiently.
+
 \section{Brushstroke Analysis}

### 2014-05-10 paper.tex
Add some conclusions.

@@
 
-This research appears to be of very limited use, but when used within as a part
-of a larger system could provide some very interesting analysis which could be
-used to train classifiers to detect forgeries, etc. but this still has limited
-scope.
-
 \section{Texture Analysis}\label{sec:texture}
@@
 pixels) at different dyadic scales ($C$), note how the picture becomes less
-pixalated at higher scales}
+pixelated at higher scales}
 \label{fig:dyadic-image}
@@
 question the validity of their results based on this.
-
-In conclusion this is a very complex but powerful technique which is actually
-applying theoretical concepts to perform some interesting analysis and
-classification. However, the sensitivity of the technique and with the need for
-expert and a posteriori decisions do limit the real application of this
-research.
 
@@
 
-In conclusion this research provides some very powerful analysis with the
-experimental results to back it up in two different fields of stylometry which
-agree with expert opinion and begin to bring forth some new ideas.
-
 \subsection{Authentication of Artwork}
@@
 
-In conclusion this work presents a method which may provide an insight into how
-many authors a work has, but has very few experimental results to back up this
-claim. Especially given other research into similar areas.
-
 % Line shaded -> applicable to paint?
@@
 
-In conclusion, this is a decent application of existing research with real
-results being gained, although it doesn't deal with the extraction of
-brushstrokes formally it is a good step in the right direction.
-
 \subsection{Rhythmic Brushstrokes}
@@
 
-In conclusion, this paper presents a very useful method for extracting
-brushstrokes and compares the process of doing so to a ground truth. There are
-still some elements which could use further investigation, but on the whole
-this is a very rounded piece of research.
-
 \subsection{Stroke Decomposition}
@@
 % TODO beast mode this.
+
+Digital analysis of artwork is not a trivial task: there are many problems
+which occur far more prevalently than in other fields of image processing. The
+need for high-resolution copies of artwork with a standard method of capture
+can significantly hinder some areas due to the lack of availability.
+
+Added to that, noise is far more common from degradation of works from a lack
+of care or poor materials used to create the piece. Although this is a
+two-sided coin and there are analysis which could be performed which look at
+specific types of degradation to authenticate an artwork.
+
+With all the different kinds of art and materials which can be used to produce
+works, it is very difficult not to over specialise in the analysis techniques,
+especially when it comes to complex techniques such as detecting brushstrokes.
+
+Texture appears the be the base of a decent analysis technique; most of the
+brushstroke extraction technique use some form of texture discrimination to
+function. Colour does provide some interesting results, but it is usually the
+cumulation of texture and colour which provides the most powerful analysis.
 

### 2014-05-10 paper.tex
Preliminary finish.
Everything is done, bib is balanced as per IEEE style reccomendations. Could do
with a read through though.

@@
 %TODO Write this once the paper is complete
+The recent trend in the digitisation of historic artwork has opened up a whole
+field of digital image processing. The digital analysis of paintings can be
+used to perform numerous tasks, from detecting forgeries to the animation of
+still life works. This work will review some of the research which has been
+performed into this area to provide a starting point for any future research.
 \end{abstract}
-
-% no keywords
 
@@
 % the document is modified later
-%\IEEEtriggeratref{8}
+\IEEEtriggeratref{17}
 % The "triggered" command can be changed if desired:
